### 2018-05-07 doc/manuscript.Rmd
expand on DAG in methods

- ALT: Alanine aminotransferase
- CV: Coefficient of variation
- DNL: De novo lipogenesis
- GEE: Generalized estimating equations
- HOMA2-%S: Homeostatic model of assessment 2 - percent sensitivity
- IGI/IR: Insulinogenic index over homeostatic model of assessment for insulin
resistance
- IS: Insulin sensitivity
- ISI: Insulin sensitivity index
- ISSI-2: Insulin Secretion-Sensitivity Index-2
- MAQ: Modified activity questionnaire
- MET: Metabolic equivalent of task
- NEFA: Non-esterified fatty acids
- OGTT: Oral glucose tolerance test
- PROMISE: Prospective Metabolism and Islet Cell Evaluation cohort
- PLS: Partial least squares
- SD: Standard deviation
- TGFA: Triacylglycerol fatty acids
- WC: Waist circumference

@@

Generalized estimating equation (GEE) models [@Zeger1986a] were used in the 
primary analysis to determine the longitudinal associations between the outcome 
variables and the predictor variables. The predictor variables
were scaled (mean-centered and standardized). Given the longitudinal 
design, an auto-regressive of order 1 working correlation matrix was 
specified in the GEE model. Covariates to adjust for were selected based on the
previous literature, from directed acyclic graph [@Greenland1999a]
recommendations, and from quasi-likelihood information criteria. 
The DAG structures to understand potential confounding, shown in `r cite_sf("dag_is")`
and `r cite_sf("dag_bcf")`, were processed by the DAGitty software
[@Textor2011a;@Shrier2008a] to generate the recommended adjustments.
The final
GEE model (M6; seen in `r cite_st('qic')`) was adjusted for years since baseline, WC, baseline age, ethnicity,
sex, ALT, MET, and total NEFA. The variables TGFA, total NEFA, sex, ethnicity, and
baseline age were classified as *time-independent* (held constant) as they were
measured only at the baseline visit or do not change throughout the study, while
the outcome variables and remaining covariates were set as *time-dependent*.
After transformations, the GEE estimates are interpreted as an expected percent 
difference in the outcome variable for every standard deviation (SD) increase in
the predictor variable given the covariates are held constant (including time). 
We also tested for an interaction with sex, ethnicity, or time by the predictor 
term for each outcome variable.

>>>>

- ALT: Alanine aminotransferase
- CV: Coefficient of variation
- DNL: De novo lipogenesis
- DAG: Directed acyclic graph
- GEE: Generalized estimating equations
- HOMA2-%S: Homeostatic model of assessment 2 - percent sensitivity
- IGI/IR: Insulinogenic index over homeostatic model of assessment for insulin
resistance
- IS: Insulin sensitivity
- ISI: Insulin sensitivity index
- ISSI-2: Insulin Secretion-Sensitivity Index-2
- MAQ: Modified activity questionnaire
- MET: Metabolic equivalent of task
- NEFA: Non-esterified fatty acids
- OGTT: Oral glucose tolerance test
- PROMISE: Prospective Metabolism and Islet Cell Evaluation cohort
- PLS: Partial least squares
- SD: Standard deviation
- TGFA: Triacylglycerol fatty acids
- WC: Waist circumference

@@

Generalized estimating equation (GEE) models [@Zeger1986a] were used in the 
primary analysis to determine the longitudinal associations between the outcome 
variables and the predictor variables. The predictor variables
were scaled (mean-centered and standardized). Given the longitudinal 
design, an auto-regressive of order 1 working correlation matrix was 
specified in the GEE model. Covariates to adjust for were selected based on the
previous literature, from directed acyclic graph (DAG) [@Greenland1999a]
recommendations, and from quasi-likelihood information criteria. 
DAGs are used to identify the minimum adjustment necessary for a model by
using the causal pathways to algorithmically identify potential confounding and
colliding variables (see the [@Greenland1999a] ref for more detail about using
DAGs). The DAG structures to understand potential confounding, shown in 
`r cite_sf("dag_is")` and `r cite_sf("dag_bcf")`, were processed by the DAGitty
software [@Textor2011a;@Shrier2008a] to generate the recommended adjustments.
These DAG structures were developed based on hypothesized causal pathways
between each variable, which were then input into the DAGitty software. The
output from DAGitty was used to help inform the final model.

The final GEE model (M6; seen in `r cite_st('qic')`) was adjusted for years
since baseline, WC, baseline age, ethnicity,
sex, ALT, MET, and total NEFA. The variables TGFA, total NEFA, sex, ethnicity, and
baseline age were classified as *time-independent* (held constant) as they were
measured only at the baseline visit or do not change throughout the study, while
the outcome variables and remaining covariates were set as *time-dependent*.
After transformations, the GEE estimates are interpreted as an expected percent 
difference in the outcome variable for every standard deviation (SD) increase in
the predictor variable given the covariates are held constant (including time). 
We also tested for an interaction with sex, ethnicity, or time by the predictor 
term for each outcome variable.
